Estimating Gaze Duration Error from Eye Tracking Data

### 2022-10-20 paper/paper.tex
Starting paper modification for precision experiment

The source repository for the gazerr application also contains a series of scripts for generating data
sets and running the experiments outlined in the next section. Full details available in the package
README file.

@@

We apply the gazerr application to investigate the relationship between measured gaze duration
and expected real duration under a variety of changing scenarios. These scenarios are variations
in the underlying error indicated by the eye tracking calibration file and length of the session
in which the user could be looking at the area of interest.

For the sake of our simulations we use a simulated device with viewport pixel 
dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium rectangle ad unit (MREC)
that is located toward the top of the screen. We iterate over a range of potential mean error
values (equivalent to varying the accuracy), determined to by the mean Euclidean distance 
between the measured fixation point and actual fixation points.
For each of these mean error values we generate a synthetic eye tracking calibration file. We produce
two versions of this calibration data, one in which the error is centered around the true target,
and a second biased version in which the error tends toward the upper left of the true position. We 
include the biased calibration data as it is consistent with our observations of real eye tracking data
and we explore the extent to which this bias affects gaze duration measurements.

We feed each of these synthetic calibration files into gazerr to calculate posterior probability
distributions over true gaze duration for the set of measured gaze durations. These posterior
distributions are converted into expected values of true gaze duration for each measured duration.
We repeat the above process for multiple session lengths to investigate the impact of increasing
session length on the expected gaze duration for a fixed set of measurements.

@@

The first experimental result is the expected true duration against the measured duration
for a single eye tracking application with a mean fixation error of 50 pixels. These
results are shown in Figure \ref{fig:measured_vs_expected}.

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
\caption{Measured versus expected gaze duration. MREC in 10s Session with 50px Mean Error.}
\label{fig:measured_vs_expected}
\end{figure}

The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we see under
multiple variations of the simulation configuration. At lower meaurement values there is a 
tendency for the measurement to be an accurate estimation, while at higher values it tends
toward an under-estimation of the true gaze duration. Surprisingly the results using the
biased calibration data seem to mitigate this effect somewhat, but the overall pattern 
remains consistent.

We next examined the way that exected error in the gaze duration estimate tracks with expected 
error in the gaze fixation points. This involved calculating the mean absolute error in gaze
duration for each of the mean fixation error validation files. We performed this calculation
using two methods. The first method, called uniform, assumes that all measured values are 
equally likely. The second method, called smoothed, assumes that low measurements are much
more likely than higher measurement, and that this drop in likelihood follows an approximately
exponential distribution.

@@

As shown in Figure \ref{fig:fixation_vs_duration} we see that the error in gaze duration
grows with error in fixation points, as would be expected. However, when we use the exponential
weighting to calculate the expected error in duration the gaze duration error grows much slower
than in the uniform case. This suggests that the observed tendency toward smaller measurements
in area of interest studies means lower expected error overall. This pattern is true for the 
centrally distributed error (A) as well as the biased calibration error (B).

>>>>

The source repository for the \textit{gazerr} application also contains a series of 
scripts for generating data sets and running the experiments outlined in the next 
section. Full details available in the package \textit{README} file.

@@

We apply the \textit{gazerr} application to investigate the relationship between 
measured gaze duration and expected true duration under a variety of changing scenarios. 
These scenarios are variations in the underlying error indicated by the eye tracking 
calibration file and length of the session in which the user could be looking at 
the area of interest.

For the sake of our experiments we use a simulated device with viewport pixel 
dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium 
rectangle ad unit (MREC) that is located toward the top of the screen. 
We iterate over a range of potential mean error values 
(equivalent to varying the accuracy), determined to by the mean Euclidean distance 
between the measured fixation point and actual fixation points.
For each of these mean error values we generate a synthetic eye tracking calibration 
file. We produce three versions of this calibration data, one in which the error 
is centered around the true target, a second biased version in which the error tends 
toward the upper left of the true position. In the third version the error is again
biased toward the upper left corner of the screen, but it has higher precision, 
meaning that the measurements will cluster closely together with each other around
the mean error position.

We include these biased calibration dataset as they are consistent with both our 
observations of real eye tracking data and what is found in the literature\cite{}.
We use this synthetic data creation in order to explore the extent to which different
patterns in error, bias and precision affects gaze duration measurements.

We feed each of these synthetic calibration files into the \textit{gazerr} application
to calculate the posterior probability distributions over true gaze duration for 
the set of measured gaze durations. These posterior distributions are converted 
into expected values of true gaze duration for each measured duration.

In these experiments we use both a fixed session length and fixed area of interest
size and position, so that we can focus specifically on the impacts of mean
error, bias and precision on the quality of gaze duration measurment.

@@

The first experimental result is the expected true duration against the measured duration
for a single eye tracking application with a mean fixation error of 70 pixels. These
results are shown in Figure \ref{fig:measured_vs_expected}.

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
\caption{Measured Versus Expected Gaze Duration. \\\hspace{\textwidth} 
MREC ad appearing on screen for 1s with 50px Mean Error in Fixation}
\label{fig:measured_vs_expected}
\end{figure}

The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we see under
multiple variations of the simulation configuration. At lower meaurement values there is a 
tendency for the measurement to be an accurate estimation of true gaze duraiton, while 
at higher values it tends toward an under-estimation of the true gaze duration. 
Surprisingly the results using the biased calibration data seem to mitigate this effect 
somewhat, but the overall pattern remains consistent.

We next examined the way that exected error in the gaze duration estimate tracks with 
expected error in the gaze fixation points. This involved calculating the mean absolute
error in gaze duration for each of the mean fixation error validation files. 
We performed this calculation using two methods. The first method, labelled uniform, 
assumes that all measured values are equally likely. The second method, labelled exponential,
assumes that low measurements are much more likely than higher measurement, and that 
this drop in likelihood follows an approximately exponential distribution.

@@

As shown in Figure \ref{fig:fixation_vs_duration} we see that the error in gaze duration
grows with error in fixation points, as would be expected. However, when we use the 
exponential weighting to calculate the expected error in duration the gaze duration 
error grows much slower than in the uniform case. This suggests that the observed 
tendency toward smaller measurements in advertising area of interest studies means 
lower expected error overall. This pattern is true for the centrally distributed 
error (A) as well as the biased calibration error (B) and precise biased calibration
error (C) scenarios.
